---
title: "Replication"
author: "Arkadzi Rakach"
date: "2026-04-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Discussion**
In this diagnostic, I tested the sensitivity of the results by splitting 
the average ideological extremity scores by ideological side. While the original 
study simply aggregated the polarization scores by taking the absolute values, 
my diagnostic seeks to maintain the structure of the original measure intact 
while adding a robustness check in the form of checking for one-sided ideological
influence. In other words, the question I ask here is: "Is the polarization 
caused mainly by one side of the political spectrum, or does it hold across the 
board?"

As can be seen below, my diagnostic graph shows that the results are not simply
driven by one side of the political aisle. Although some social media topics 
such as minimum wage and Syria have slight, albeit notable, difference in 
polarization between liberals and conservatives, the reality is that the results 
prove to be quite robust.

After this diagnostic, I am slightly surprised at the decision of the authors 
to not include this in the paper for this figure, seeing that it might have
strengthened their arguments. They do test this in the poisson regression they
run, but doing it for this figure would have been much more intuitive for the
average reader.

Finally, this diagnostic is valid for this article due to the fact that it
analyzes networks of retweets on certain issues, assigned with an ideology
score. Although my diagnostic does not ensure that the ideology assignment is
correct, it does make sure that the results are not dependent on a single 
ideology, building confidence in the study.

**Limitations**
One major limitation is that I was not able to include data from the 2012 
election. I went to the data lab in order to see why the data for this was not
loading, and it turned out that the authors had missing data for the election
on their GitHub.

The data lab helped me figure out the diagnostic code and problem with 
missing 2012 election data. I will attach photos showing the clearly incorrect
file size and empty file which is available on their GitHub to the assignment
submission on Canvas. 

In the future, it may be interesting to look at how polarization on
different sides of the political aisle differs by political and 
non-political topics.


```{r}
load("output/estimates.rdata")
```

```{r}
library(reshape)
library(ggplot2)
library(scales)
library(RColorBrewer)
```

**3C**
```{r}
lists <- c("minimum_wage", "obama", "superbowl", "marriageequality", "budget",
           "newtown", "governmentshutdown", "oscars", "sotu", "syria",
           "olympics", "boston")

labels <- c("Minimum Wage", "2012 Election", "Super Bowl", "Marriage Equality",
            "Budget", "Newtown Shooting", "Govt. Shutdown", "Oscars 2014",
            "State of the Union", "Syria", "Winter Olympics", "Boston Marathon")

polarization <- c()

for (j in 1:length(lists)){
  
  cat(j, " ")
  
  lst <- lists[j]
  cat(lst, "\n")
  
  load(paste0("output/temp/retweets_ideology/rt_results_", lst, ".rdata"))
  
  polarization <- rbind(
    polarization,
    data.frame(
      collection = labels[j],
      polar = mean(abs(results$ideology_retweeted), na.rm = TRUE),
      stringsAsFactors = FALSE
    )
  )
}

polarization <- polarization[order(polarization$polar), ]

polarization$collection <- factor(
  polarization$collection,
  levels = rev(polarization$collection)
)

p <- ggplot(polarization, aes(x = collection, y = polar, fill = polar, label = collection))

pq <- p +
  geom_bar(stat = "identity") +
  theme_bw() +
  coord_flip() +
  scale_y_continuous(
    "Aggregate Ideological Polarization",
    expand = c(0, 0),
    limits = c(0, 4.0)
  ) +
  geom_text(hjust = 1.05, color = "white", size = 3) +
  theme(
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )

pq

ggsave(filename = "plots/figure3c.pdf", plot = pq, height = 4, width = 3)
```


**3C Diagnostic**

This is where I separate liberal and conservative retweets by individual issues. 

```{r}
lists <- c("minimum_wage", "superbowl", "marriageequality", "budget",
           "newtown", "governmentshutdown", "oscars", "sotu", "syria",
           "olympics", "boston")

labels <- c("Minimum Wage", "Super Bowl", "Marriage Equality",
            "Budget", "Newtown Shooting", "Govt. Shutdown", "Oscars 2014",
            "State of the Union", "Syria", "Winter Olympics", "Boston Marathon")

polarization_liberal <- c()
polarization_conservative <- c()

for (j in 1:length(lists)){
  
  cat(j, " ")
  
  lst <- lists[j]
  cat(lst, "\n")
  
  load(paste0("output/temp/retweets_ideology/rt_results_", lst, ".rdata"))
  
  liberal_results <- results[results$ideology_retweeted < 0, ]
  conservative_results <- results[results$ideology_retweeted > 0, ]
  
  polarization_liberal <- rbind(
    polarization_liberal,
    data.frame(
      collection = labels[j],
      polar = mean(abs(liberal_results$ideology_retweeted), na.rm = TRUE),
      side = "Liberal content",
      n = nrow(liberal_results),
      stringsAsFactors = FALSE
    )
  )
  
  polarization_conservative <- rbind(
    polarization_conservative,
    data.frame(
      collection = labels[j],
      polar = mean(abs(conservative_results$ideology_retweeted), na.rm = TRUE),
      side = "Conservative content",
      n = nrow(conservative_results),
      stringsAsFactors = FALSE
    )
  )
}

polarization_signed <- rbind(polarization_liberal, polarization_conservative)

polarization_all <- c()

for (j in 1:length(lists)){
  
  lst <- lists[j]
  
  load(paste0("output/temp/retweets_ideology/rt_results_", lst, ".rdata"))
  
  polarization_all <- rbind(
    polarization_all,
    data.frame(
      collection = labels[j],
      polar_all = mean(abs(results$ideology_retweeted), na.rm = TRUE),
      stringsAsFactors = FALSE
    )
  )
}

polarization_all <- polarization_all[order(polarization_all$polar_all), ]

polarization_signed$collection <- factor(
  polarization_signed$collection,
  levels = polarization_all$collection
)

pq_diag <- ggplot(
  polarization_signed,
  aes(x = collection, y = polar, fill = side)
) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_bw() +
  coord_flip() +
  scale_y_continuous(
    "Aggregate Ideological Polarization",
    expand = c(0, 0),
    limits = c(0, 4.0)
  ) +
  labs(
    title = "Diagnostic: Left vs. Right Retweet Polarization",
    x = NULL,
    fill = "Retweeted author ideology"
  )

pq_diag

ggsave(filename = "plots/figure3c_diagnostic.pdf",
       plot = pq_diag, height = 4, width = 6)

```




