---
title: "Extension"
author: "Arkadzi Rakach"
date: "2026-06-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Introduction**
The article by Barberá et al. (2015) studies whether political communication 
over the social media platform of Twitter functions as an echo chamber or if 
there is notable broader cross-ideological engagement. To do this, they estimate 
the ideological positions of Twitter users and then examine 
retweeting across twelve political and nonpolitical topics. The authors
classify topics such as the 2012 election, government shutdown, budget, marriage
equality, minimum wage, and the State of the Union as political topics. The 
topics were not manually classified by the content of each tweet individually, 
but rather, the authors assigned tweets to topics using keyword searches. Tweets
were assigned to a collection if they mentioned one of the relevant search terms
for that topic, such as “shutdown” or “boehner” for the government shutdown, or
“Syria,” “intervention,” and “chemical weapons” for Syria. They classify topics 
such as the Boston Marathon bombing, Newtown shooting, Oscars, Super Bowl, 
Winter Olympics, and Syria as nonpolitical topics. Their general finding is that
political topics tend to produce more ideologically polarized retweeting, while 
many nonpolitical topics have more cross-ideological retweeting.


```{r}
load("output/estimates.rdata")
```

```{r}
library(reshape)
library(ggplot2)
library(scales)
library(RColorBrewer)
```


```{r}
load("output/estimates.rdata")
estimates <- estimates[,c("uid", "ideology")]
```


**Replicating Heat Maps from Figures 2A and 2B**

The first step in this process is replicating the heat maps themselves. This is 
relatively straightforward, taking the already created retweet polarization
results per each topic included in the study. With the exception of the 
changing the file destinations slightly, the code ran as expected. Once again, 
however, the data for the 2012 election is missing, which appears to be a 
problem with the Harvard Dataverse page itself. 


```{r}
lists <- c("minimum_wage", "superbowl", "marriageequality", "budget",
  "newtown", "governmentshutdown", "oscars", "sotu", "syria",
  "olympics", "boston")

not.poli <- c("superbowl", "newtown", "oscars", "syria", "olympics", "boston")
not.poli.labels <- c("Super Bowl", "Newtown Shooting", "Oscars 2014", "Syria",
    "Winter Olympics", "Boston Marathon")
poli <- c("minimum_wage", "marriageequality", "budget", "governmentshutdown",
    "sotu")
poli.labels <- c("Minimum Wage", "Marriage Equality", "Budget",
    "Gov. Shutdown", "State of the Union")

## functions to construct heatmaps
min <- -3.5
max <- 3.5
breaks <- 0.25

expand_data <- function(breaks=0.10, min=-4, max=4){
    x <- results$ideology_retweeter
    y <- results$ideology_retweeted
    x <- (round((x - min) / breaks, 0) * breaks) + min
    y <- (round((y - min) / breaks, 0) * breaks) + min
    tab <- table(x, y)
    tab <- melt(tab)
    tab$prop <- tab$value/sum(tab$value)
    return(tab)
}
```


**Figure 2A**

First, the political topics heat map is created. Again, please note the absence
of the 2012 election.


```{r}
xy.poli <- c()

for (j in 1:length(poli)){

    cat(j, " ")

    lst <- poli[j]
    cat(lst, " ")

    load(paste0("output/temp/retweets_ideology/rt_results_", lst, ".rdata"))

    new.xy <- expand_data(breaks = 0.25)
    new.xy$candidate <- poli.labels[j]
    xy.poli <- rbind(xy.poli, new.xy)
}


p <- ggplot(xy.poli, aes(x = y, y = x))
pq <- p + geom_tile(aes(fill = prop), colour = "white") + 
        scale_fill_gradient(name = "% of\ntweets", 
        low = "white", high = "black", 
        breaks = c(0, .0050, 0.010, 0.015, 0.02), limits = c(0, .021),
        labels = c("0.0%", "0.5%", "1.0%", "1.5%", ">2%")) +
        labs(y = "Estimated Ideology of Retweeter", x = "Estimated Ideology of Author") + 
        scale_y_continuous(expand = c(0,0), breaks = (-2:2), limits = c(-3, 3)) +
        scale_x_continuous(expand = c(0,0), breaks = (-2:2), limits = c(-3, 3)) +
        facet_wrap(~ candidate, nrow = 2) + 
        theme(panel.border = element_rect(fill = NA), panel.background = element_blank()) +
        coord_equal() 

pq

ggsave(filename = "plots/figure2a.pdf", plot = pq, height = 5, width = 8)
```

**Figure 2B**

The process is then repeated for non-political topics. 


```{r}
xy <- c()

for (j in 1:length(not.poli)){

    cat(j, " ")

    lst <- not.poli[j]
    cat(lst, " ")

    load(paste0("output/temp/retweets_ideology/rt_results_", lst, ".rdata"))

    new.xy <- expand_data(breaks = 0.25)
    new.xy$candidate <- not.poli.labels[j]
    xy <- rbind(xy, new.xy)
}

p <- ggplot(xy, aes(x = y, y = x))
pq <- p + geom_tile(aes(fill = prop), colour = "white") + 
        scale_fill_gradient(name = "% of\ntweets", 
        low = "white", high = "black", 
        breaks = c(0, .0050, 0.010, 0.015, 0.02), limits = c(0, .021),
        labels = c("0.0%", "0.5%", "1.0%", "1.5%", ">2%")) +
        labs(y = "Estimated Ideology of Retweeter", x = "Estimated Ideology of Author") + 
        scale_y_continuous(expand = c(0,0), breaks = (-2:2), limits = c(-3, 3)) +
        scale_x_continuous(expand = c(0,0), breaks = (-2:2), limits = c(-3, 3)) +
        facet_wrap(~ candidate, nrow = 2) + 
        theme(panel.border = element_rect(fill = NA), panel.background = element_blank()) +
        coord_equal()

pq

ggsave(filename = "plots/figure2b.pdf", plot = pq, height = 5, width = 8)
```


**Extension 1**

I decided to run an extension in the form of a simple robustness check. First,
I wanted to run an aggregate polarization map for all political and
non-political topics. Doing so allows for a clearer category-level visualization 
of the overall retweet structure across political and nonpolitical topics. 
Rather than comparing each topic separately, the pooled heatmaps show whether
the broader political and nonpolitical distinction remains visible when retweets 
are aggregated within each category.

First, I assign all the retweet results data to the objects of "pol" and 
"non_pol" in the same way they are assigned in the original article. 

```{r}
pol <- c("minimum_wage", "marriageequality", "budget", 
         "governmentshutdown", "sotu")

non_pol <- c("superbowl", "newtown", "oscars", 
             "syria", "olympics", "boston")
```

```{r}
political_results <- c()

for (topic in pol) {
  
  load(paste0("output/temp/retweets_ideology/rt_results_", topic, ".rdata"))
  
  political_results <- rbind(political_results, results)
}
```

```{r}
nonpolitical_results <- c()

for (topic in non_pol) {
  
  load(paste0("output/temp/retweets_ideology/rt_results_", topic, ".rdata"))
  
  nonpolitical_results <- rbind(nonpolitical_results, results)
}
```

```{r}
results <- political_results
xy.pol.total <- expand_data(breaks = 0.25)
xy.pol.total$type <- "Political"

results <- nonpolitical_results
xy.nonpol.total <- expand_data(breaks = 0.25)
xy.nonpol.total$type <- "Nonpolitical"
```


```{r}
xy.total <- rbind(xy.pol.total, xy.nonpol.total)
```


```{r}
p <- ggplot(xy.total, aes(x = y, y = x))

pq <- p + 
  geom_tile(aes(fill = prop), colour = "white") + 
  scale_fill_gradient(
    name = "% of\ntweets", 
    low = "white", high = "black", 
    breaks = c(0, .0050, 0.010, 0.015, 0.02),
    limits = c(0, .021),
    labels = c("0.0%", "0.5%", "1.0%", "1.5%", ">2%")
  ) +
  labs(
    y = "Estimated Ideology of Retweeter",
    x = "Estimated Ideology of Author"
  ) + 
  scale_y_continuous(expand = c(0,0), breaks = (-2:2), limits = c(-3, 3)) +
  scale_x_continuous(expand = c(0,0), breaks = (-2:2), limits = c(-3, 3)) +
  facet_wrap(~ type, nrow = 1) + 
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_blank()
  ) +
  coord_equal()

pq
```


**Extension 2**

I decided to run a robustness check by assigning "Syria" to the political 
topics. Since Syria is a topic which can very plausibly be considered 
as a political one, I wanted to test whether the polarization results were
dependent on arbitrary, or perhaps convenient assignment of categories.

The same process is repeated here, but some objects are renamed to avoid 
overwriting ones from the previous heat maps. 


```{r}
pol2 <- c("minimum_wage", "marriageequality", "budget", 
         "governmentshutdown", "sotu", "syria")

non_pol2 <- c("superbowl", "newtown", "oscars", 
             "olympics", "boston")
```

```{r}
political_results2 <- c()

for (topic in pol2) {
  
  load(paste0("output/temp/retweets_ideology/rt_results_", topic, ".rdata"))
  
  political_results2 <- rbind(political_results2, results)
}
```

```{r}
nonpolitical_results2 <- c()

for (topic in non_pol2) {
  
  load(paste0("output/temp/retweets_ideology/rt_results_", topic, ".rdata"))
  
  nonpolitical_results2 <- rbind(nonpolitical_results2, results)
}
```

```{r}
results <- political_results2
xy.pol.total2 <- expand_data(breaks = 0.25)
xy.pol.total2$type <- "Political, Syria Included"

results <- nonpolitical_results2
xy.nonpol.total2 <- expand_data(breaks = 0.25)
xy.nonpol.total2$type <- "Nonpolitical, Syria Excluded"
```

```{r}
xy.total2 <- rbind(xy.pol.total2, xy.nonpol.total2)
```

```{r}
p <- ggplot(xy.total2, aes(x = y, y = x))

pq <- p + 
  geom_tile(aes(fill = prop), colour = "white") + 
  scale_fill_gradient(
    name = "% of\ntweets", 
    low = "white", high = "black", 
    breaks = c(0, .0050, 0.010, 0.015, 0.02),
    limits = c(0, .021),
    labels = c("0.0%", "0.5%", "1.0%", "1.5%", ">2%")
  ) +
  labs(
    y = "Estimated Ideology of Retweeter",
    x = "Estimated Ideology of Author"
  ) + 
  scale_y_continuous(expand = c(0,0), breaks = (-2:2), limits = c(-3, 3)) +
  scale_x_continuous(expand = c(0,0), breaks = (-2:2), limits = c(-3, 3)) +
  facet_wrap(~ type, nrow = 1) + 
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_blank()
  ) +
  coord_equal()

pq
```


**Discussion**

This extension is appropriate because it remains close to the method used in 
the original paper while adding a natural extension in the form of a robustness 
check. Indeed, this allows one to examine whether the political and nonpolitical 
divide remains visible at the broader category label rather than just within 
topics. The combined heat maps demonstrate that both the political and 
nonpolitical topics support the findings from the original article. 
Although political topics generate more ideologically polarized retweets, the 
retweets from nonpolitical topics are generally more cross-ideological.

The robustness test using the change in Syria classification demonstrates that 
reassigning retweets on the topic of Syria from the nonpolitical to the 
political category does not significantly change the aggregate findings. In 
fact, it only further strengthens the findings in the original article, as 
reassigning the category of Syria actually appears to decrease polarized 
retweets in the top right quadrant.


**Limitations**

One of the limitations regarding this extension is that topics with
significantly more retweets may sway the results of the combined heat maps.
At the same time, aggregated political heat map does not include the most 
tweeted topic of the 2012 election, which based on the original paper, is also 
one of the most polarized topics. This means that it would have potentially 
heavily influenced the results here if it was included. 

Moreover, there are still values which are dropped from both heat maps due to
ideological values being outside of the -3 to 3 range.
