---
title: "Replication"
author: "Arkadzi Rakach"
date: "2026-04-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Discussion**
In this replication, I replicated figures 3C and 3D from the article "Tweeting From Left to Right: Is Online
Political Communication More Than an Echo Chamber?" Their code was all available on github and Harvard 
Dataverse, so I was able to replicate it without many  issues. However, I ran into an issue on figure 4C, where 
two rows do not show up in my graph. I also made some minor changes and adjustments on the environment-level, in 
terms of document locations, to the original code.


```{r}
install.packages("reshape")
install.packages("scales")
```


```{r}
load("output/estimates.rdata")
```

```{r}
library(reshape)
library(ggplot2)
library(scales)
```


```{r}
lists <- c("minimum_wage", "obama", "superbowl", "marriageequality", "budget",
           "newtown", "governmentshutdown", "oscars", "sotu", "syria",
           "olympics", "boston")

labels <- c("Minimum Wage", "2012 Election", "Super Bowl", "Marriage Equality",
            "Budget", "Newtown Shooting", "Govt. Shutdown", "Oscars 2014",
            "State of the Union", "Syria", "Winter Olympics", "Boston Marathon")

polarization <- c()

for (j in 1:length(lists)){
  
  cat(j, " ")
  
  lst <- lists[j]
  cat(lst, "\n")
  
  load(paste0("output/temp/retweets_ideology/rt_results_", lst, ".rdata"))
  
  polarization <- rbind(
    polarization,
    data.frame(
      collection = labels[j],
      polar = mean(abs(results$ideology_retweeted), na.rm = TRUE),
      stringsAsFactors = FALSE
    )
  )
}

polarization <- polarization[order(polarization$polar), ]

polarization$collection <- factor(
  polarization$collection,
  levels = rev(polarization$collection)
)

p <- ggplot(polarization, aes(x = collection, y = polar, fill = polar, label = collection))

pq <- p +
  geom_bar(stat = "identity") +
  theme_bw() +
  coord_flip() +
  scale_y_continuous(
    "Aggregate Ideological Polarization",
    expand = c(0, 0),
    limits = c(0, 1.7)
  ) +
  geom_text(hjust = 1.05, color = "white", size = 3) +
  theme(
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )

pq

ggsave(filename = "plots/figure3c.pdf", plot = pq, height = 4, width = 3)
```


```{r}
install.packages("RColorBrewer")
```

```{r}
library(RColorBrewer)
```


```{r}
load("output/temp/polarization-by-day.rdata")

plot.data <- pol.days[pol.days$collection %in% c("2012 Election",
    "Marriage Equality", "Newtown Shooting", "Syria", "Boston Marathon"),]
plot.data <- plot.data[plot.data$day<30,]

plot.data$collection <- factor(plot.data$collection, levels=c("2012 Election",
    "Marriage Equality", "Newtown Shooting", "Syria", "Boston Marathon"))

colors <- c(
  rev(brewer.pal(5, "RdBu"))[1],
  rev(brewer.pal(5, "RdBu"))[5],
  "purple",
  "green",
  "orange"
)

p <- ggplot(plot.data, aes(x=day, y=polar, group=collection))
pq <- p + geom_point(aes(color=collection), alpha=1/2, size=1.5) + stat_smooth(aes(color=collection)) +
    scale_x_continuous(name="Days of Twitter Data Collection", limits=c(0,30), expand=c(0,0)) +
      theme_bw() + 
    scale_y_continuous("Polarization", limits=c(0.50,2.00), expand=c(0,0)) +
    theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.key.size=unit(0.5, "cm"),
    legend.key.height=unit(2,"line")) +
    scale_color_manual("Collection", breaks=c("2012 Election",
    "Marriage Equality", "Newtown Shooting", "Syria", "Boston Marathon"), 
        labels=c("2012\nElection",
    "Marriage\nEquality", "Newtown\nShooting", "Syria", "Boston\nMarathon"), values=colors)
pq

ggsave(filename = "plots/figure3d.pdf", plot = pq, height = 3.5, width = 6)
```

